---
output:
  md_document:
    variant: markdown_github
---


# PresElectionResults

```{r include=FALSE}
dataraw_dir <- "/home/jtimm/Dropbox/GitHub/packages/PresElectionResults/data-raw/"
data_dir <- '/home/jtimm/Dropbox/GitHub/packages/PresElectionResults/data/'
```


```{r message=FALSE, warning=FALSE}
library(dplyr)
```


## Presidential Election Results -- Britannica

https://www.britannica.com/topic/United-States-Presidential-Election-Results-1788863

> Sources: Electoral and popular vote totals based on data from the Office of the Clerk of the U.S. House of Representatives; the United States Office of the Federal Register; the Federal Election Commission; Congressional Quarterly's Guide to U.S. Elections, 4th ed. (2001); and the official certified state vote totals.

```{r message=FALSE, warning=FALSE}
url99 <- 'https://www.britannica.com/topic/United-States-Presidential-Election-Results-1788863'

# Try multiple selectors to find the table
page <- url99 |> xml2::read_html()

# Try different XPath selectors in order of specificity
ecs_node <- rvest::html_node(page, xpath = '//*[@class="md-raw-html"]/div/table/tbody')

# If that doesn't work, try without tbody
if (inherits(ecs_node, "xml_missing")) {
  ecs_node <- rvest::html_node(page, xpath = '//*[@class="md-raw-html"]/div/table')
}

if (inherits(ecs_node, "xml_missing")) {
  ecs_node <- rvest::html_node(page, xpath = '//table/tbody')
}

if (inherits(ecs_node, "xml_missing")) {
  # Try finding all tables and take the first one (likely the main table)
  all_tables <- rvest::html_nodes(page, css = 'table')
  if (length(all_tables) > 0) {
    ecs_node <- all_tables[[1]]
  }
}

if (inherits(ecs_node, "xml_missing")) {
  stop("Could not find the presidential election results table on Britannica page. The page structure may have changed. Please inspect the page manually.")
}

ecs <- rvest::html_table(ecs_node, fill = TRUE) 

colnames(ecs) <- c('year', 
                   'candidate', 
                   'party', 
                   'ec_votes', 
                   'pop_votes', 
                   'pop_per')


fns <- data.frame(year = c(1789,
                           1800, 
                           1800,
                           1824,
                           2000),
                  
                  candidate = c('George Washington', 
                                'Thomas Jefferson', 
                                'Aaron Burr',
                                'John Quincy Adams',
                                'Al Gore'),
                  z = 1)


pres_results <- ecs |>
  mutate(pop_votes = as.integer(gsub(',', '', pop_votes)),
         pop_per = as.numeric(pop_per),
         candidate = gsub('[0-9]', '', candidate) |> trimws(),
         candidate = gsub('George Bush$', 'George H.W. Bush', candidate),
         
         # Horace Greeley
         party = gsub('/Liberal Republican', '', party),
         party = gsub('[0-9]', '', party),
         party = gsub('no distinct party designations|no formally organized parties', 
                                '', 
                                party),
         party = trimws(party)) |>

  left_join(fns) |>
  mutate(ec_votes = ifelse(!is.na(z), 
                           gsub('.$', '', ec_votes), 
                           ec_votes),
         ec_votes = ifelse(candidate == 'Al Gore', 
                           gsub('.$', '', ec_votes), 
                           ec_votes),
         
         ec_votes = as.integer(ec_votes)) |>
  select(-z) |>
  
  group_by(year) |>
  mutate(ec_total = sum(ec_votes, na.rm = T)) |> ungroup() |>
  filter(!party %in% 'not a candidate',
         !candidate %in% 'not voted')
```





## Presidential election results by county (2000-2024) -- MIT Election Data and Science Lab (MEDSL)

https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/VOQCHQ

**NOTE: The MEDSL county data has known quality issues.** Kansas 2024 county-level FIPS codes appear to be scrambled -- votes are present but assigned to the wrong counties, producing implausible margin swings of 30-40pp. Similar issues may exist in other states/years and are not easily detectable without an independent source to cross-check against. For county-level analysis, consider [tonmcg/US_County_Level_Election_Results_08-24](https://github.com/tonmcg/US_County_Level_Election_Results_08-24) as an alternative.

The county build is retained below for reference but should be used with caution.

```{r message=FALSE, warning=FALSE, eval=FALSE}
library(dplyr)
setwd(dataraw_dir)
county <- read.delim('countypres_2000-2024.tab', sep = '\t')

pres_by_county <- county |>
  filter(!is.na(candidatevotes),
         !candidate %in% c("TOTAL VOTES CAST")) |>
  group_by(year, state_po, county_name, county_fips, candidate, party) |>
  mutate(has_total = any(mode %in% c("TOTAL", "TOTAL VOTES"))) |>
  filter(!has_total | mode %in% c("TOTAL", "TOTAL VOTES")) |>
  ungroup() |>
  group_by_at(vars(all_of(colnames(county)[c(1:8,10)]))) |>
  summarize(candidatevotes = sum(candidatevotes),
            totalvotes     = mean(totalvotes)) |> ungroup() |>
  group_by(year, state_po, county_name, county_fips) |>
  mutate(totalvotes_actual = sum(candidatevotes)) |>
  ungroup() |>
  mutate(per = round(candidatevotes / totalvotes_actual * 100, 1)) |>
  mutate(party = tolower(party),
         county_name = stringr::str_to_title(tolower(county_name))) |>
  group_by(year, state_po, county_name, county_fips) |>
  mutate(winner =  candidate[which.max(candidatevotes)],
         winner = stringr::str_to_title(tolower(winner)),
         party_win = party[which.max(candidatevotes)]) |>
  ungroup() |>
  select(-candidate, -candidatevotes) |>
  filter(party %in% c('republican', 'democrat')) |>
  tidyr::spread(party, per) |>
  rename(state_abbrev = state_po) |>
  mutate(GEOID = stringr::str_pad(county_fips, 5, pad = "0")) |>
  select(year, state_abbrev, county_name, GEOID, winner, party_win, democrat, republican)
```




### Extract house rep names from 1976-2024-house.tab

```{r message=FALSE, warning=FALSE}
# Read house election data to extract winning house reps
setwd(dataraw_dir)
house_data <- read.delim('1976-2024-house.tab', sep = ',', stringsAsFactors = FALSE) |>
  janitor::clean_names() |>
  filter(year == 2024) |>
  
  filter(office == 'US HOUSE',
         stage == 'GEN',
         writein == FALSE) |>
  group_by(year, state_po, district) |>
  # Get winner (highest vote getter) for each year/state/district
  mutate(winner_votes = max(candidatevotes, na.rm = TRUE)) |>
  filter(candidatevotes == winner_votes) |>
  # Handle ties by taking first (should be rare)
  slice(1) |>
  ungroup() |>
  mutate(
    state_abbrev = state_po,
    district_code = ifelse(district == 'AL', '00', 
                           stringr::str_pad(as.character(district), width = 2, side = 'left', pad = '0')),
    house_rep = stringr::str_to_title(tolower(candidate)),
    house_rep_party = tolower(party)
  ) |>
  mutate(
    house_rep_party = ifelse(house_rep_party == 'democratic-farmer-labor', 'democrat', house_rep_party)
  ) |>
  select(year, state_abbrev, district_code, house_rep, house_rep_party) |>
  # Create last name and normalized name for ICPSR matching
  mutate(
    last = gsub('^.* ', '', house_rep) |> toupper(),
    nn_kos = house_rep |> tolower() |> trimws()
  ) |>
  group_by(year, last) |>
  mutate(last_n = n()) |>
  ungroup() 
```



### 2024 Presidential results by CD

https://docs.google.com/spreadsheets/d/1ZHx5E0-5vuXxcZShBgsAl_vwAntkkanGqYQp0owNjoQ/edit?gid=0#gid=0

```{r}
setwd(dataraw_dir)
pres_by_cd00 <- read.csv('2024 Pres by CD - Main.csv', check.names = FALSE) |> 
  janitor::clean_names() |>
  mutate(
    # Extract percentages from Harris % and Trump % columns (remove % sign)
    harris_pct = as.numeric(gsub('%', '', harris_percent)),
    trump_pct = as.numeric(gsub('%', '', trump_percent)),
    winner = ifelse(winner == 'Harris', 
                    'Kamala Harris', 
                    'Donald Trump'),
    party_win = ifelse(harris_pct > trump_pct, 
                       'democrat', 
                       'republican'),
    # Add NA for house rep info since not in 2024 file
    house_rep = NA_character_,
    house_rep_party = NA_character_
  ) |> 
  
  rename(democrat = harris_pct,
         republican = trump_pct) |>
  
  tidyr::separate(cd, into = c('state_abbrev', 'district_code'), sep = '-') |>
  mutate(district_code = gsub('AL', '00', district_code)) |>
  mutate(year = 2024) |>
  # Merge with house rep data from 1976-2024-house.tab
  left_join(house_data |> filter(year == 2024) |> 
              select(year, state_abbrev, district_code, house_rep, house_rep_party, last, nn_kos, last_n),
            by = c('year', 'state_abbrev', 'district_code')) |>
  # Use house rep from join if available (house_rep.x and house_rep_party.x are NA from initial mutate)
  mutate(
    house_rep = coalesce(house_rep.y, house_rep.x),
    house_rep_party = coalesce(house_rep_party.y, house_rep_party.x),
    # last, nn_kos, last_n only come from house_data (no .x versions, so they keep their names)
    # Handle NAs for last_n
    last_n = coalesce(last_n, 0L)
  ) |>
  select(state_abbrev, district_code, 
         house_rep, house_rep_party,
         winner, party_win, 
         democrat, republican,
         last, nn_kos, last_n) |>
  ungroup()
```




### Add ICPSR info from VoteView

```{r message=FALSE, warning=FALSE}
# Download house member metadata for 118th Congress (2024)
hm1 <- Rvoteview::download_metadata(type = 'members',
                                    chamber = 'house',
                                    congress = '119') |>
  mutate(bioname = stringi::stri_trans_general(bioname, id = "Latin-ASCII"),
         nn_vv = gsub(', Jr\\.', '', bioname),
         nn_vv = gsub('(.*)(, )(.*$)', '\\3 \\1', nn_vv),
         nn_vv = stringr::str_to_title(nn_vv), 
         nn_vv = gsub(' [A-Z]\\.', '', nn_vv) |> tolower() |> trimws(),
         last = gsub(',.*$', '', bioname) |> trimws()) |>
  group_by(last) |>
  mutate(last_n = n()) |>
  ungroup()

# Match house reps to ICPSR codes (only if house_rep is available)
pbd <- pres_by_cd00 |>
  filter(!is.na(last)) |>
  left_join(hm1 |> filter(last_n == 1) |> select(nn_vv, last),
            by = 'last') |>
  select(nn_kos, nn_vv)

kos <- pbd |> filter(is.na(nn_vv))

strReverse <- function(x) sapply(lapply(strsplit(x, NULL), rev), paste, collapse="")

get_closest <- function(x, y) {
  y[stringdist::stringdist(x |> strReverse(), 
                         y |> strReverse(),
                         method = "jw") |>  which.min()]
}

pbd0 <- pbd |>
  rowwise() |>
  mutate(nn_vv = ifelse(is.na(nn_vv),
                        get_closest(x = nn_kos, y = hm1$nn_vv),
                        nn_vv))

pres_by_cd <- pres_by_cd00 |>
  left_join(pbd0 |> select(nn_kos, nn_vv)) |>
  mutate(nn_vv = ifelse(nn_kos == 'chris smith', 'christopher henry smith', nn_vv),
         nn_vv = ifelse(nn_kos == 'sanford bishop', 'sanford dixon bishop', nn_vv),
         nn_vv = ifelse(nn_kos == 'hal rogers', 'harold dallas (hal) rogers', nn_vv)) |>
  left_join(hm1 |> select(nn_vv, icpsr)) |>
  inner_join(pres_by_cd00) |>
  select(state_abbrev, district_code, 
         house_rep, house_rep_party,
         winner, party_win, 
         democrat, republican, icpsr)
```


## Presidential election results by state (1864-) -- Wikipedia

Eg:

https://en.wikipedia.org/wiki/United_States_presidential_elections_in_Virginia

### Table details by state

```{r message=FALSE, warning=FALSE}
options(tigris_use_cache = TRUE, tigris_class = "sf")
states_full <- tigris::states(cb = TRUE) |> 
  data.frame() |>
  select(NAME, STATEFP, STUSPS) |>
  rename(state_abbrev = STUSPS)

###
w3a <- c('New Mexico', 
         'Alaska', 
         'Arizona', 
         'District of Columbia',
         'Hawaii',
         'Connecticut',
         'New York',
         'Washington',
         'Oklahoma')  # Connecticut, New York, Washington, and Oklahoma use same column structure as w3a
w3c <- c('Pennsylvania')
w3d <- c('California')
w5 <- c('Nebraska')
w4a <- c('Maine')
w4b <- c('Utah', 'Wyoming')
w6 <- c('Arkansas', 'Florida')
w7 <- c('Wisconsin')  # Wisconsin has 21 columns, needs special handling

states <- states_full |>
  
  mutate (          l1 = case_when (NAME %in% c(w3a, w6, w4b, w5) ~ list(list(1,3,4,5)), # Nebraska: Year (col1), Winner candidate (col3), votes (col4), % (col5)
                          !NAME %in% c(w3a, w3c, w3d, w6, w4b, w5, w7) ~ list(list(1,2,3,4)),
                          NAME %in% w3c ~ list(list(1,5,6,6)),
                          NAME %in% w3d ~ list(list(1,2,5,6)),
                          NAME %in% w7 ~ list(list(1,3,4,5))),
          
          l2 = case_when (NAME %in% c(w3a, w6, w4b) ~ list(list(1,7,8, 9)),
                          NAME %in% w5 ~ list(list(1,8,9,10)),  # Nebraska: Year (col1), Runner-up candidate (col8), votes (col9), % (col10)
                          !NAME %in% c(w3a, w3c, w3d, w6, w4b, w5, w7) ~ list(list(1,5,6,7)),
                          NAME %in% w3c ~ list(list(1,9,10,10)),
                          NAME %in% w3d ~ list(list(1,8,11,12)),
                          NAME %in% w7 ~ list(list(1,7,8,9))),
          
          l3 = case_when (NAME == 'Connecticut' ~ list(list(1,10,11,12)),  # Connecticut: Year (col1), Other candidate (col10), votes (col11), % (col12)
                          NAME == 'New York' ~ list(list(1,10,11,12)),  # New York: Year (col1), Other candidate (col10), votes (col11), % (col12)
                          NAME == 'Washington' ~ list(list(1,10,11,12)),  # Washington: Year (col1), Other candidate (col10), votes (col11), % (col12)
                          NAME == 'Oklahoma' ~ list(list(1,10,11,12)),  # Oklahoma: Year (col1), Other candidate (col10), votes (col11), % (col12)
                          NAME %in% c(w3a, w6, w4b) ~ list(list(1,11, 12, 13)),
                          NAME %in% w5 ~ list(list(1,12,13,14)),  # Nebraska: Year (col1), Third candidate (col12), votes (col13), % (col14)
                          !NAME %in% c(w3a, w3c, w3d, w6, w4b, w5, w7, 'Connecticut', 'New York', 'Washington', 'Oklahoma') ~ list(list(1,8, 9, 10)),
                          NAME %in% w3c ~ list(list(1,1,1,1)),
                          NAME %in% w3d ~ list(list(1,1,1,1)),
                          NAME %in% w7 ~ list(list(1,10,11,12))),
          
          ns = case_when (NAME == 'Connecticut' ~ 7,  # Connecticut uses table 7 - must come BEFORE w3a check
                          NAME == 'New York' ~ 8,  # New York uses table 8 - must come BEFORE w3a check
                          NAME == 'Washington' ~ 3,  # Washington uses table 3 - must come BEFORE w3a check
                          NAME == 'Oklahoma' ~ 3,  # Oklahoma uses table 3 - same structure as Washington
                          NAME %in% c(w3a, w3c) ~ 3,  # New Mexico group and Pennsylvania (Connecticut, New York, Washington excluded by above)
                          NAME %in% w3d ~ 3,   # California uses table 3 (updated page structure)
                          NAME %in% w7 ~ 14,  # Wisconsin uses table 14 (third wikitable, 1864-present only)
                          NAME %in% c(w4a, w4b) ~ 4,
                          NAME %in% c(w5) ~ 3,  # Nebraska uses table 3
                          NAME %in% c(w6) ~ 6,
                          !NAME %in% c(w3a, w3c, w3d, w4a, w4b, w5, w6, w7, 'Connecticut', 'New York', 'Washington', 'Oklahoma') ~ 2)
  ) |>
  
  mutate(ns = ns) |>
  filter(!STATEFP %in% c('78', '69', '66', '72', '60')) |>
  mutate(NAME = gsub('Washington', 'Washington (state)', NAME))
```



### Scrape tables

```{r}
base_url <- 
  'https://en.wikipedia.org/wiki/United_States_presidential_elections_in_'
```



```{r message=FALSE, warning=FALSE}
states_correct <- list()

for (i in 1:nrow(states)) {
  
  state_url <- paste0(base_url, gsub(' ', '_', states$NAME[i]))
  state_page <- xml2::read_html(state_url)
  
  # Try primary selector first
  table_node <- rvest::html_node(
    state_page,
    xpath = paste0('//*[@id="mw-content-text"]/div/table[', 
                   states$ns[i],']'))
  
  # If that fails, try fallback selectors
  if (inherits(table_node, "xml_missing")) {
    # Try alternative XPath without specific ID
    table_node <- rvest::html_node(
      state_page,
      xpath = paste0('//*[@id="mw-content-text"]/table[', 
                     states$ns[i],']'))
  }
  
  if (inherits(table_node, "xml_missing")) {
    # Try finding all tables and take the ns-th one
    all_tables <- rvest::html_nodes(state_page, xpath = '//*[@id="mw-content-text"]//table')
    if (length(all_tables) >= states$ns[i]) {
      table_node <- all_tables[[states$ns[i]]]
    }
  }
  
  # Check if we found a table before trying to parse it
  if (inherits(table_node, "xml_missing")) {
    warning(paste("Could not find table for state:", states$NAME[i], 
                  "at table index", states$ns[i]))
    states_correct[[i]] <- data.frame(year = integer(), 
                                      candidate = character(),
                                      votes = character(),  # character to match html_table output
                                      vote_share = character(),
                                      stringsAsFactors = FALSE)
    next
  }
  
  # Try to parse the table
  tryCatch({
    states_correct[[i]] <- rvest::html_table(table_node, fill = TRUE)
    
    if(nrow(states_correct[[i]]) > 2){
      
      # Extract columns - handle errors if columns don't exist
      l1_vals <- states$l1[i] |> unlist()
      l2_vals <- states$l2[i] |> unlist()
      
      # Check if we have enough columns in the table
      max_col_l1 <- max(l1_vals)
      max_col_l2 <- max(l2_vals)
      
      if(max_col_l1 > ncol(states_correct[[i]]) || max_col_l2 > ncol(states_correct[[i]])) {
        warning(paste("Not enough columns for", states$NAME[i], "- skipping"))
        states_correct[[i]] <- data.frame(year = integer(), candidate = character(),
                                          votes = character(), vote_share = character(),
                                          stringsAsFactors = FALSE)
        next
      }
      
      # Extract l1 and l2 columns - these should always exist
      x <- NULL
      y <- NULL
      
      tryCatch({
        x <- states_correct[[i]][, l1_vals, drop = FALSE]
        colnames(x) <- c('year', 'candidate', 'votes', 'vote_share')
        y <- states_correct[[i]][, l2_vals, drop = FALSE]
        colnames(y) <- c('year', 'candidate', 'votes', 'vote_share')
      }, error = function(e) {
        warning(paste("Error extracting l1/l2 columns for", states$NAME[i], "-", e$message))
        states_correct[[i]] <<- data.frame(year = integer(), candidate = character(),
                                          votes = character(), vote_share = character(),
                                          stringsAsFactors = FALSE)
      })
      
      # If extraction failed, skip to next state
      if(is.null(x) || is.null(y)) {
        next
      }
      
      # California's l3 is broken (1,1,1,1) - skip it
      # Also skip l3 if not enough columns in the table (some rows may not have "Other" candidates)
      l3_vals <- states$l3[i] |> unlist()
      max_col_needed <- max(l3_vals)
      
      if(all(l3_vals == c(1,1,1,1))) {
        states_correct[[i]] <- rbind(x, y)
      } else if(max_col_needed > ncol(states_correct[[i]])) {
        # Not enough columns for l3, skip it
        states_correct[[i]] <- rbind(x, y)
      } else {
        # Try to extract l3, but handle errors if columns don't exist
        tryCatch({
          z <- states_correct[[i]][, l3_vals, drop = FALSE]
          colnames(z) <- c('year', 'candidate', 'votes', 'vote_share')
          states_correct[[i]] <- rbind(x, y, z)
url99 <- 'https://www.britannica.com/topic/United-States-Presidential-Election-Results-1788863'
        }, error = function(e) {
          # If l3 extraction fails, just use x and y
          states_correct[[i]] <<- rbind(x, y)
        })
      }
      
      states_correct[[i]] <- states_correct[[i]] |>
        mutate(
          year = gsub("\\D+", "", year),
          year = as.integer(substr(year, 1,4)),
          # Handle cases where votes and percentage are combined (e.g., "86,110(56.58%)")
          # Extract just the numeric part for votes
          votes = gsub('\\(.*\\)|%', '', votes) |> trimws(),
          vote_share = as.character(vote_share)
        ) |>
        # Filter out obviously invalid years (before 1788 or after 2024)
        filter(year >= 1788, year <= 2024, !is.na(year))
    }
  }, error = function(e) {
    warning(paste("Error processing table for state:", states$NAME[i], 
                  "-", e$message))
    states_correct[[i]] <<- data.frame(year = integer(), 
                                      candidate = character(),
                                      votes = character(),  # character to match html_table output
                                      vote_share = character(),
                                      stringsAsFactors = FALSE)
  })
} 

names(states_correct) <- states$NAME
states_correct1 <- states_correct |>
  bind_rows(.id = 'state_name')  |>
  mutate(votes = as.integer(gsub('\\(.*\\)|,', '', votes))) |>
  mutate(vote_share = gsub('(^.*\\()(.*)(\\)$)', '\\2', vote_share),
         vote_share =as.numeric(gsub('%', '', vote_share)),
         # wiki_party = gsub('(^.*\\()([A-Za-z.]*)(\\)$)', '\\2', candidate),
         candidate = gsub('\\(.*\\)|\\[.*\\]| ‡', '', candidate) |> trimws()) |>

  filter(candidate != 'TBD',
         nchar(candidate) > 5,
         !is.na(year)) |>
  
  # Filter out metadata rows - these have newlines, election type names, or year lists in candidate
  filter(!grepl("\n", candidate)) |>  # Candidate names don't have newlines
  filter(!grepl("^(Presidential|Gubernatorial|State propositions|Mayoral|County|Board of|General|U\\.S\\.|Elections in|Secretary of|House of|Governor)", candidate, ignore.case = TRUE)) |>
  filter(!grepl("^(Dem|Rep|Democratic|Republican) [0-9]{4}", candidate, ignore.case = TRUE)) |>
  filter(!grepl("^[0-9]{4}–[0-9]{4}$", candidate)) |>  # Year ranges like "1910–1919"
  filter(!(grepl("^[A-Z][a-z]+$", candidate) & nchar(candidate) < 15)) |>  # Single word city/place names
  filter(!is.na(votes) & votes != "") |>  # Must have votes data
  
  ## correct hubert humphrey in florida 1972
  mutate(candidate = ifelse(state_name == 'Florida' & year == 1972 & candidate == 'Hubert Humphrey',
                            'George McGovern', candidate))
```







### Resolve wiki and britannica candidate names

```{r}
brits <- unique(pres_results$candidate)
wikis <- unique(states_correct1$candidate)
## setdiff(brits, wikis)

nnb <- sapply(wikis, function(x){
   brits[adist(x, brits) |> which.min()]
})

xx <- data.frame(wiki = nnb |> names() |> trimws(),
                 brit = nnb |> unname() |> trimws()) |>
  unique() 

ms <- c('Adlai Stevenson II',
        'Al Smith',
        'Charles Pinckney')

cross <- xx |>
  # mutate(brit = ifelse(wiki %in% ms, NA, brit)) |>
  mutate(wiki =  ifelse(brit == 'William H. Crawford', 'William H. Crawford', wiki)) |>
  mutate(brit = ifelse(wiki == 'John Q. Adams', 'John Quincy Adams', brit)) |> 
  mutate(brit = ifelse(wiki == 'Charles Pinckney', 'Charles Cotesworth Pinckney', brit)) |> 
  unique() |> na.omit()
```




### Final election return table

```{r message=FALSE, warning=FALSE}
pres_by_state <- states_correct1 |>
  mutate(vote_share = ifelse(year %in% c('1864', '1868') &
                               vote_share == 0,
                             NA, vote_share),
         
         candidate = trimws(candidate)) |>
  
  ## na.omit() |> 
  left_join(cross, by = c('candidate' = 'wiki')) |>
  
  left_join(pres_results |> select(year:party) |>
              mutate(party = ifelse(party %in% 'Democratic', 'Democrat', party)), 
            by = c('brit' = 'candidate',
                                                'year' = 'year')) |>
  mutate(party = ifelse(candidate %in% c('Al Smith', 'Adlai Stevenson II'),
                        'Democrat', party)) |>
  
  rename(NAME = state_name) |>
  left_join(states) |>
  select(year, state_abbrev, brit, party, votes:vote_share) |>
  rename(candidate = brit) |>
  filter(year >= 1864, !is.na(vote_share)) |>
  select(-votes) |>
  unique() |>
  mutate(party = ifelse(!party %in% c('Democrat', 'Republican'), 
                        'Other', party)) |>
  
  group_by(year, state_abbrev, party) |>
  filter(vote_share == max(vote_share)) |>
  ungroup() |>
  
  group_by(state_abbrev, year) |>   
  mutate(winner = candidate[which.max(vote_share)],
         #party = tolower(party),
         party_win = tolower(party[which.max(vote_share)])) |>
  ungroup() |>
  
  select(-candidate) |>
  
  unique() |>
  tidyr::spread(party, vote_share)

colnames(pres_by_state) <- colnames(pres_by_state) |> tolower() 
```
  



## Equal-area simple feature geometries -- The Downballot

https://docs.google.com/spreadsheets/d/13XkF59JKzvw4SeSq5mbgIFrJfYjK4amg9JoQE5e9grQ/edit?gid=0#gid=0

```{r}
fnames <- c('HexCDv30wm',
            'HexSTv30wm',
            'TileInv10', 
            'TileOutv10')

sfs <- lapply(fnames, function(x) {

  sf::st_read(dsn = paste0(dataraw_dir, x),
                     layer = x,
                     quiet = TRUE) })
names(sfs) <- fnames

xsf_HexCDv30wm <- sfs$HexCDv30 |>
  rename(state_abbrev = STATEAB,
         state = STATENAME) |>
  mutate(district_code = gsub('[A-Z][A-Z]', 0, CDLABEL),
         district_code = stringr::str_pad (district_code, 2, pad = 0)) |>
  select(GEOID, state, state_abbrev, district_code)


xsf_HexSTv30wm <- sfs$HexSTv30wm |>
  rename(state_abbrev = STATEAB, state = STATENAME)
    
xsf_TileInv10 <- sfs$TileInv10 %>% select(5:7) %>%
  rename(state_abbrev = State, state = StateName)

xsf_TileOutv10 <- sfs$TileOutv10 %>% select(5:7) %>%
  rename(state_abbrev = State, state = StateName)


xsf_HexCDv30wm |> head()
```


## FRED Historical Population data

https://fred.stlouisfed.org/release/tables?rid=118&eid=259194&od=#

```{r message=FALSE, warning=FALSE}
freds <- lapply(1900:2026, function(x){
  
  ur <- paste0('https://fred.stlouisfed.org/release/tables?rid=118&eid=259194&od=', x, '-01-01#')
  
  ur |>
   xml2::read_html() |>
    rvest::html_node(
      xpath = '//*[@class="table-responsive"]/table/tbody') |>
    rvest::html_table(fill = TRUE) |>
    
    mutate(year = x,
           NAME = X2) |>
    
    rowwise() |>
    mutate(population = ifelse(x < 1950, 
                               gsub('\\.|,', '', X4) |> as.integer(),
                               gsub('\\.|,', '', X3) |> as.integer())) |>
    ungroup() |>
    select(year, NAME, population)
  
})

fred_pop_by_state <- freds |>
  bind_rows() |>
  #filter(!is.na(population)) |>
  left_join(states_full) |>
  select(1,5,2, 3)

fred_pop_by_state |>
  head()|>
  knitr::kable()
```


## Output

```{r message=FALSE, warning=FALSE, include=TRUE}
setwd(data_dir)
usethis::use_data(pres_by_state, overwrite=TRUE)
usethis::use_data(pres_by_cd, overwrite=TRUE)
#usethis::use_data(pres_by_county, overwrite=TRUE)
usethis::use_data(pres_results, overwrite=TRUE)

usethis::use_data(fred_pop_by_state, overwrite=TRUE)

# usethis::use_data(xsf_HexCDv30wm, overwrite=TRUE)
# usethis::use_data(xsf_HexSTv30wm, overwrite=TRUE)
usethis::use_data(xsf_TileInv10, overwrite=TRUE)
usethis::use_data(xsf_TileOutv10, overwrite=TRUE)

# if(exists('xsf_ElectoralCollege')) {
#   usethis::use_data(xsf_ElectoralCollege, overwrite=TRUE)
# }
```
