---
output:
  md_document:
    variant: markdown_github
---


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path =  "figure-markdown_github/")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("/home/jtimm/pCloudDrive/GitHub/packages/render_toc.R")
```


# PresElectionResults

A brief reference package to US Presidential Elections.

```{r echo=FALSE}
render_toc("/home/jtimm/pCloudDrive/GitHub/packages/PresElectionResults/README.Rmd",
           toc_header_name = "PresElectionResults",
           #base_level = NULL,
           toc_depth = 3)
```


## Installation

```{r message=FALSE, warning=FALSE}
library(dplyr)
```


```{r eval=FALSE}
library(devtools)
devtools::install_github("jaytimm/PresElectionResults")
```



## Presidential Election Results via Britannica

```{r}
PresElectionResults::pres_results |>
  tail() |> knitr::kable()
```


## Presidential election results by county (2000-2020) via MIT Election Data and Science Lab (MEDSL)

```{r}
PresElectionResults::pres_by_county |>
  head() |> knitr::kable()
```


## Presidential election results by congressional district (2020) via Daily Kos

```{r}
PresElectionResults::pres_by_cd |>
  head() |> knitr::kable()
```


## Presidential election results by state (1864-) via Wikipedia

```{r}
PresElectionResults::pres_by_state |>
  head() |> knitr::kable()
```



## Equal-area simple feature geometries via Daily Kos

```{r fig.height=6.5, message=FALSE, warning=FALSE}
# devtools::install_github("yutannihilation/ggsflabel")
library(sf)
library(ggplot2)
PresElectionResults::xsf_HexCDv30wm |> 
  ggplot() + 
  geom_sf(aes(fill = state),
          color = 'white') +

  ggsflabel::geom_sf_text(data = PresElectionResults::xsf_HexSTv30wm,
                          aes(label = state_abbrev),
                          size = 3,
                          color='black') +
  
  ## scale_fill_distiller(palette = "RdBu", direction=-1) +
  theme_minimal()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position = 'none') +
labs(title = "Equal-area US Congressional District geometry",
     caption = "Source: DailyKos")
```



```{r fig.height=6.5, message=FALSE, warning=FALSE}
library(sf)
library(ggplot2)
PresElectionResults::xsf_TileOutv10 %>% 
  ggplot() + 
  geom_sf(aes(fill = state),
           color = 'white') +
  geom_sf(data = PresElectionResults::xsf_TileInv10, 
          fill = NA, 
          show.legend = F, 
          color="gray", 
          lwd=.5) +
  ggsflabel::geom_sf_text(data = PresElectionResults::xsf_TileInv10,
                          aes(label = state_abbrev),
                          size = 3,
                          color='black') +
  ##scale_fill_distiller(palette = "RdBu", direction=-1) +
  theme_minimal()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position = 'none') +
labs(title = "Equal-area US State geometry",
     caption = "Source: DailyKos")
```







