---
output:
  md_document:
    variant: markdown_github
---


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path =  "figure-markdown_github/")
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("/home/jtimm/jt_work/GitHub/git_projects/render_toc.R")
```


# uspols

A collection of US political data, including (1) federal election returns, (2) Twitter details for US lawmakers, and (3) some relevant political boundary files as simple feature geometries.  Collated from existing resources, eg, [MEDSL](https://github.com/MEDSL) and [DailyKos](https://www.dailykos.com/stories/2018/2/21/1742660/-The-ultimate-Daily-Kos-Elections-guide-to-all-of-our-data-sets#1), and included here as an **R data package**.  


```{r echo=FALSE}
render_toc("/home/jtimm/jt_work/GitHub/packages/uspols/README.Rmd",
           toc_header_name = "uspols",
           #base_level = NULL,
           toc_depth = 3)
```


Formats have been tweaked for uniform output across data sets.  Links to R code demonstrate all details of work-flow from raw data to package table.     

## Installation

```{r message=FALSE, warning=FALSE}
library(tidyverse)
```


```{r eval=FALSE}
library(devtools)
devtools::install_github("jaytimm/uspols")
library(uspols) 
```


## Details 

### MEDSL - House returns by congressional district - 1976-2018

[Code](https://github.com/jaytimm/uspols/blob/master/mds/medsl.md) 

```{r message=FALSE, warning=FALSE}
uspols::uspols_medsl_house %>%
  head() %>% knitr::kable()

```

### MEDSL - Senate returns by state - 1976-2018

```{r eval=FALSE}
uspols::uspols_medsl_senate
```

---

### DailyKos - Presidential returns by congressional district - 2008-2016

[Code](https://github.com/jaytimm/uspols/blob/master/mds/daily-kos.md) | [DailyKos dataset](https://docs.google.com/spreadsheets/d/1zLNAuRqPauss00HDz4XbTH2HqsCzMe0pR8QmD1K8jk8/edit#gid=0)

```{r message=FALSE, warning=FALSE}
uspols::uspols_dk_pres %>%
  head() %>% knitr::kable() 
```


---

### Wikipedia - Presidential returns by state - 1864-2016

[Code](https://github.com/jaytimm/uspols/blob/master/mds/wikipedia.md)

```{r message=FALSE, warning=FALSE}
uspols::uspols_wiki_pres %>%
  head() %>% knitr::kable()
```


---

### MEDSL - Presidential returns by county - 2000-2016

[Code](https://github.com/jaytimm/uspols/blob/master/mds/medsl-county.md)

```{r}
uspols::medsl_pres_county %>% head() 
```


---

### Lawmaker Twitter handles - Congresses 115 and 116

[Code](https://github.com/jaytimm/twitter-and-us-lawmakers/blob/master/twitter-handles.md)

```{r}
uspols::uspols_twitter_handles %>% head() %>% knitr::kable()
```


---


### DailyKos - Simple feature geometries

[States shapefile](https://docs.google.com/spreadsheets/d/1LrBXlqrtSZwyYOkpEEXFwQggvtR0bHHTxs9kq4kjOjw/edit#gid=1278876419)

```{r eval=FALSE, fig.height=6.5, message=FALSE, warning=FALSE, include=FALSE}
#[Congressional Districts](https://docs.google.com/spreadsheets/d/1LrBXlqrtSZwyYOkpEEXFwQggvtR0bHHTxs9kq4kjOjw/edit#gid=1250379179)
library(sf)
uspols::sf_HexCDv21 %>%
  ggplot() + 
  geom_sf(aes(fill = state),
           color = 'white') +
  geom_sf(data = uspols::sf_HexSTv20, 
          fill = NA, 
          show.legend = F, 
          color="gray", 
          lwd=.5) +
  ggsflabel::geom_sf_text(data = uspols::sf_HexCDv21,
                           aes(label = district_code), 
                           size = 2,
                           color='black') +
  # scale_fill_distiller(palette = "RdBu", direction=-1) +
  theme_minimal()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position = 'none') +
labs(title = "Equal-area US Congressional District geometry",
     caption = "Source: DailyKos")
```



```{r fig.height=6.5, message=FALSE, warning=FALSE}
library(sf)
uspols::xsf_TileOutv10 %>% 
  ggplot() + 
  geom_sf(aes(fill = state),
           color = 'white') +
  geom_sf(data = uspols::xsf_TileInv10, 
          fill = NA, 
          show.legend = F, 
          color="gray", 
          lwd=.5) +
  ggsflabel::geom_sf_text(data = uspols::xsf_TileInv10,
                          aes(label = state_abbrev), 
                          size = 3,
                          color='black') +
  # scale_fill_distiller(palette = "RdBu", direction=-1) +
  theme_minimal()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position = 'none') +
labs(title = "Equal-area US State geometry",
     caption = "Source: DailyKos")
```


---


### Wikipedia - Timeline of the Trump presidency

`uspols_wiki_timeline()`: a simple function for scraping a Wikipedia-based [timeline of the Trump presidency](https://en.wikipedia.org/wiki/Timeline_of_the_Donald_Trump_presidency).  The function returns an up-to-date table of daily administration happenings from 20 Jan 2017 onward.   Event data have been structured for a uniform output.

```{r}
timeline <- uspols::uspols_wiki_timeline() 

timeline %>%
  select(-weekof, -daypres, -dow) %>%
  filter(date == '2017-01-20') %>%
  knitr::kable() ## Day 1 -- 20 Jan 2017 --
```







