---
output:
  md_document:
    variant: markdown_github
---


```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "##",
  fig.path =  "figure-markdown_github/")
```


# PresElectionResults

A data package for US Presidential election results.  

> All build details are available [here](https://github.com/jaytimm/PresElectionResults/blob/master/builds.md), and are fully reproducible.


## Installation

```{r message=FALSE, warning=FALSE}
library(dplyr)
```


```{r eval=FALSE}
devtools::install_github("jaytimm/PresElectionResults")
```



## Presidential Election Results via Britannica

```{r}
PresElectionResults::pres_results |>
  tail() |> knitr::kable()
```


## Presidential election results by state (1864-2024) via Wikipedia

```{r}
PresElectionResults::pres_by_state |>
  head() |> knitr::kable()
```



## Presidential election results by congressional district (2024)

```{r}
PresElectionResults::pres_by_cd |>
  head() |> knitr::kable()
```



## Equal-area simple feature geometries via The Downballot

```{r fig.height=6.5, message=FALSE, warning=FALSE}
library(sf)
library(ggplot2)
PresElectionResults::xsf_TileOutv10 %>% 
  ggplot() + 
  geom_sf(aes(fill = state),
           color = 'white') +
  geom_sf(data = PresElectionResults::xsf_TileInv10, 
          fill = NA, 
          show.legend = F, 
          color="gray", 
          lwd=.5) +
  
  geom_sf_text(data = PresElectionResults::xsf_TileInv10,
                          aes(label = state_abbrev),
                          size = 3,
                          color='black') +
  ##scale_fill_distiller(palette = "RdBu", direction=-1) +
  theme_minimal()+
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank(),
        legend.position = 'none') +
labs(title = "Equal-area US State geometry",
     caption = "Source: The Downballot")
```




## FRED Historical Population Data (1900-2026)

```{r}
PresElectionResults::fred_pop_by_state |>
  head()|>
  knitr::kable()
```



